---
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
Impact of Weather Events on the Economy and Population Health in U.S.A 
===================================================================================

# Synopsis 
Weather events, such as tornadoes, storms and flood have caused a lot of trouble on public health and economic consequences. 

This report will show what type of the weather events have the most impact on population and Economy in U.S.A. 

_The data was collected from 1950 to Novenber 2011._  


# Data Processing 

## Reading Data ##

- Data is downloaded from:
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2
- The name of data flie is "repdata-data-StormData.csv.bz2".

```{r cache=TRUE}
Sys.setlocale('LC_ALL', 'English') #If your system is not in English. 
stormdata=read.csv("repdata-data-StormData.csv.bz2")
```

## Data Transformations ##


### Time from 1950 to Nov. 2011

```{r echo=TRUE}
stormdata$BGN_DATE=as.Date(stormdata$BGN_DATE, format = "%m/%d/%Y")
stormdata.after1950= stormdata[(stormdata$BGN_DATE >= "1950-01-01"), ]
str(stormdata.after1950$BGN_DATE)
```

Thus, we have 902297 observations and 37 variables:

```{r echo=TRUE}
dim(stormdata.after1950)
```

### Loading required packages 
```{r echo=TRUE}
library(Hmisc)
library(reshape)
library(car)
library(lattice)
library(ggplot2)

``` 

### Regulating the events' types(EVTYPE) 

Change the EVTYPE into small letter:
```{r echo=TRUE}
stormdata$EVTYPE=capitalize(tolower(stormdata$EVTYPE))
``` 

The Data is then processed and analyzed to:
- The effect on *population health*: "pop.health" dataframe which is summrized by fatalities and injuries. 
- The effect on *economy*: "eco.dam" dataframe which is summrized by properity damage and crop damage.
- "pop.health.top" and "eco.dam.top"" will show top 10 events.


### The effect on *population health*:
```{r echo=TRUE}
pop.health=aggregate(cbind(FATALITIES, INJURIES) ~ EVTYPE, stormdata, sum)
```

Top 10 for each event:
```{r echo=TRUE}
ph1=head(pop.health[order(-pop.health$FATALITIES),c(1,2) ], 10)
ph2=head(pop.health[order(-pop.health$INJURIES),c(1,3) ], 10)
ph1
ph2
```
Top 10 for the sum:
```{r echo=TRUE}
ph.sum=aggregate(FATALITIES+INJURIES ~ EVTYPE, stormdata, sum)
head(ph.sum[order(-ph.sum[,2]),c(1,2) ], 10) 
```
### The effect on *economy*:

First, we convert PROPDMG*PROPDMGEXP and CROPDMG*CROPDMGEXP into the format of "dollar"
```{r echo=TRUE}
stormdata$PROPDMG <- stormdata$PROPDMG * as.numeric(Recode(stormdata$PROPDMGEXP, 
    "'0'=1;'1'=10;'2'=100;'3'=1000;'4'=10000;'5'=100000;'6'=1000000;'7'=10000000;'8'=100000000;'B'=1000000000;'h'=100;'H'=100;'K'=1000;'m'=1000000;'M'=1000000;'-'=0;'?'=0;'+'=0", 
    as.factor.result = FALSE))
stormdata$CROPDMG <- stormdata$CROPDMG * as.numeric(Recode(stormdata$CROPDMGEXP, 
    "'0'=1;'2'=100;'B'=1000000000;'k'=1000;'K'=1000;'m'=1000000;'M'=1000000;''=0;'?'=0", 
    as.factor.result = FALSE))
```

Then, we have effect and top10:

```{r echo=TRUE}
eco.dam=aggregate(cbind(PROPDMG, CROPDMG) ~ EVTYPE, stormdata, sum)
```

Top 10 for each event:
```{r echo=TRUE}
head(eco.dam[order(-eco.dam$PROPDMG),c(1,2) ], 10)
head(eco.dam[order(-eco.dam$CROPDMG),c(1,3) ], 10)
```
Top 10 for the sum:
```{r echo=TRUE}
eco.sum=aggregate(PROPDMG+CROPDMG ~ EVTYPE, stormdata, sum)
head(eco.sum[order(-eco.sum[,2]),c(1,2) ], 10)
```

# Results

## Population Health:
### Question: 
#### Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?


```{r echo=TRUE}
pop.health.top.melt=melt(head(pop.health[order(-pop.health$FATALITIES, -pop.health$INJURIES), ], 10))
ggplot(pop.health.top.melt, aes(x = EVTYPE, y = value, fill = variable)) + geom_bar(stat = "identity") + 
    coord_flip() + ggtitle("Population Health") + labs(x = "", y = "number of people") + 
    scale_fill_manual(values = c("black", "green"), labels = c("Fatalities", "Injuries"))
```


## Economic Consequences
### Question:
#### Across the United States, which types of events have the greatest economic consequences?

```{r echo=TRUE}
eco.dam.top.melt=melt(head(eco.dam[order(-eco.dam$PROPDMG, -eco.dam$CROPDMG), ], 10))
ggplot(eco.dam.top.melt, aes(x = EVTYPE, y = value, fill = variable)) + geom_bar(stat = "identity") + 
    coord_flip() + ggtitle("Economic consequences") + labs(x = "", y = "cost in dollars") + 
    scale_fill_manual(values = c("red", "black"), labels = c("Property Damage", "Crop Damage"))
```

## From the above plots, we can conclude that: 
- The weather event which has the most impact on *population health* is **Tornado**.
- The weather event which has the most impact on *economic consequences* is **Flood**.


